---
title: "Homework 5"
author: "Tucker Morgan (tlm2152)"
date: "11/20/2021"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
```


## Problem 1

```{r Homicide Raw Data, message = FALSE}
hom_url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

hom_data <- 
  read_csv(hom_url, na = c("", "Unknown"))
```

The homicide data set above, `hom_data`, from _The Washington Post_ includes `r nrow(hom_data)` observations of `r ncol(hom_data)` variables including `city`, `state`, victim characteristics, and `disposition` characterizing the outcome or status of the homicide.

```{r Homicide city_state and summaries}
city_state_summ <- 
  hom_data %>% 
  unite(city_state, c("city", "state"), sep = ", ") %>% 
  filter(city_state != "Tulsa, AL") %>% 
  group_by(city_state) %>% 
  summarise(total_hom = n(),
            total_unsolved = sum(disposition == "Closed without arrest" | disposition == "Open/No arrest"))
```

```{r Baltimore prop test}
balt_md <- 
  city_state_summ %>% 
  filter(city_state == "Baltimore, MD") 

balt_prop <- 
  prop.test(x = balt_md %>% pull(total_unsolved),
            n = balt_md %>% pull(total_hom)) %>% 
  broom::tidy() %>% 
  mutate(conf.low = round(conf.low, digits = 4)) %>% 
  mutate(conf.high = round(conf.high, digits = 4)) %>% 
  select(estimate, conf.low, conf.high)
```

Now I'll create a function to iterate across each city, which mimics the process above.
```{r iterating function}
prop_est_function <- function(city){
  
  city_df <- 
    city_state_summ %>% 
    filter(city_state == city)
  
  city_prop <- 
    prop.test(x = city_df %>% pull(total_unsolved),
              n = city_df %>% pull(total_hom))
  
  output = broom::tidy(city_prop)
  
}
```

Here, I use `map()` to iterate the function across each `city_state`.
```{r city state output, warning = FALSE}
homicide_est <- 
  city_state_summ %>%  
  mutate(prop_est = map(.x = city_state, .f = prop_est_function)) %>% 
  unnest(prop_est) %>% 
  select(city_state, total_hom, total_unsolved, estimate, conf.low, conf.high) %>% 
  janitor::clean_names()
```

Below, I create a plot to show the point estimates and confidence intervals for proportion of unsolved homicides in each city.
```{r city state plot}
homicide_est %>% 
  mutate(city_state = fct_reorder(city_state, estimate, .desc = FALSE)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high)) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = -90, vjust = .35, hjust = 0),
        plot.title = element_text(face = "bold")) +
  labs(x = "City", y = "Estimated Proportion of Unsolved Homicides", title = "Estimates of Unsolved Homicide Proportions in 50 US Cities")
```


## Problem 2


Below, I will start by reading in a list of file names from the `study_data` sub-directory.
```{r study data import}
study_files <- 
  tibble(file_name = list.files(path = "./study_data/")) %>% 
  mutate(file_name = str_c("./study_data/", file_name))
```

I added a string containing the file path to the `file_name` column temporarily to facilitate the function below.

```{r study data function}
read_study_data = function(filename){
  
  file = read_csv(filename) %>% 
    pivot_longer(cols = week_1:week_8,
                 names_to = "study_week",
                 names_prefix = "week_",
                 values_to = "study_obs")
  
  output <- file
}
```

In this function, R will read in each data file and clean up the data frame by pivoting the `week_#` variable into values for `study_week` and `study_obs`. Next, I will iterate the function over each file using the `file_name` column in my data frame above.

```{r iteration, message = FALSE}
study_df <- 
  study_files %>% 
  mutate(subj_data = map(.x = file_name, .f = read_study_data)) %>% 
  unnest(cols = subj_data)

head(study_df)
```

This iterative function read in the data for each `.csv` file and cleaned the resulting data frames as described above. To view the full results, I used `unnest()`. Next, I will clean up the `file_name` column and convert it into columns for study arm and subject ID.

```{r cleaning study data}
study_df <- 
  study_df %>% 
  mutate(file_name = str_replace(file_name, ".csv", "")) %>% 
  mutate(file_name = str_replace(file_name, "./study_data/", "")) %>% 
  separate(file_name,
           into = c("study_arm", "subj_num"),
           sep = "_",
           remove = FALSE) %>% 
  rename(subj_id = file_name) %>% 
  select(-subj_num)

head(study_df)
```

```{r study spaghetti plot}
study_df %>% 
  group_by(subj_id) %>% 
  ggplot(aes(x = study_week, y = study_obs, group = subj_id, color = study_arm)) +
  geom_line() +
  labs(x = "Study Week", y = "Observed Measurement", title = "Control vs. Experimental Arm", color = "Study Arm") +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

Above, I have produced a spaghetti plot showing the observations for each `subj_id` over time, `study_week`. Each line is colored according to the subject's `study_arm`. It seems that the experimental arm saw greater values for the study observations over time compared to the control arm.

## Problem 3

First, I'll load in the `iris` dataset and introduce some missing values.
```{r making iris missing}
iris_with_missing <- 
  iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

The function below will check that the input is a character or numeric value, then it will replace character NA values with "virginica" and replace numeric NA values with the mean of the provided vector.
```{r replacement function}
iris_replace <- function(x) {
  
    if (is.character(x)) {
        x[is.na(x)] = "virginica"
    } else if (is.numeric(x)) {
        x[is.na(x)] = mean(x, na.rm = TRUE)
    } else stop("Input must be a numeric or character value.")
  return(x)
}
```

I'll use a `for` loop to iterate across each column.
```{r iris replacement}
  for (i in 1:5) {
    iris_with_missing[i] = map(iris_with_missing[i], ~iris_replace(.x))
  }
```

The below data frame should include any rows with NA values.
```{r}
na_check <- iris_with_missing[rowSums(is.na(iris_with_missing)) > 0,]
na_check
```

Looks like everything has been replaced properly.
